\Large \bf SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing

### 2016-03-05 papers/2016-HotStorage/paper.tex
shifted format to USENIX style and fixed all refs

\documentclass{sig-alt-gov2}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}

\newcommand{\DDT}{D\textsuperscript{2}T~}
\newcommand{\DDTns}{D\textsuperscript{2}T}

\hyphenation{sub-trans-ac-tion}

\addtolength{\parskip}{-0.08in}

% in order for balance columns to work, it has to be before begin document...
%\balancecolumns

\begin{document}

%\conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
%\CopyrightYear{2015}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

\title{SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}

\numberofauthors{1}
\author{
%\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2],Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
\alignauthor {Jay Lofstead\footnotemark[1], Carlos Maltzahn\footnotemark[2], Scott Klasky\footnotemark[3], Hasan Abbasi\footnotemark[3], Qing Liu\footnotemark[3]\\
Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
\footnotemark[1]~~Sandia National Laboratories
\footnotemark[2]~~University of California\\
\footnotemark[3]~~Oak Ridge National Laboratory
\footnotemark[4]~~Rutgers University
\footnotemark[5]~~Brown University\\
}
\maketitle

\begin{abstract}

The longest existing parallel I/O APIs, HDF5, PnetCDF, and NetCDF, are focused
on treating the entire output from a paralell application as a single unit.
These libraries work to gather the data from all processes and arrange it such
that it is organized much like as if it were written from a single process.
This approach, in particular the two-phase collective I/O with data sieving
portion, has proven difficult to scale for 3-D domain decompositions at scale.
ADIOS shifted the focus from treating the entire application output as a single
entity to treating the output from each writer as a self-contained entity. This
eliminated the need for the problematic two-phase collective I/O and has
demonstrated good scalability well into the petascale era. Another shift is
required to address the needs of exascale systems. New system I/O bandwidth is
not keeping pace with the compute acceleration leading to accute bottlenecks.
While ``burst buffer'' technology will help address the performance gap, it
does not address the problem completely. This paper describes a new approach
to not just an I/O library, but also the underlying storage infrastructure to
address the needs of exascacle applications cognizant of projected exascale
platform characteristics and in the context of current industry trends.

\end{abstract}

%\category{D.4}{Software}{Operating Systems}
%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

%\terms{Design, Performance}

@@

The first generation of standardized APIs for scientific computing IO focused
on serial output from a single process to a distributed or local file system.
These systems still exist in the form of serial HDF5~\cite{hdf5} and the
end-of-life NetCDF 3.x~\cite{netcdf}. The key feature of these APIs was the
logical data model that eliminated any parallel data decomoposition artifacts
from the actual data storage. This offered consistent performance and regular
data access patterns.

The second generation is really an evolution of the first with HDF-5 adding
parallel capabilities and NetCDF bifurcating into NetCDF4~\cite{netcdf4} and
PnetCDF~\cite{pnetcdf} with PnetCDF almost completely maintaining the data
model and NetCDF4 shifting to use HDF5 underneath. The other improvement is
incorporating knowledge of an underlying parallel file system to support large
files written in unision from many processes.

The third generation is a reaction to the second based on recognizing that
maintaining the logical data model in physical form became inordinately
expensive and offered a performance penalty in most cases when running at
scale. Libraries like ADIOS~\cite{adios} shifted from writing complete
variables from distributed processes as a single entity to treating all data
from a single source to be treated as a unit. This eliminated the data
reorganization at the cost of more complicated reading patterns. In spite of
this extra complication, the performance is generally
better~\cite{six-degrees}.

@@

\noindent\textbf{Refactoring} may be a uniform operation across an
entire variable. For example, splitting a double into three parts--two byes,
four bytes, and two bytes, yields a two-byte value that is the majority of the
data precision. The next four bytes would be most of the rest of the data
precision. The last two bytes are the remainder of the precision. For quick
overviews, just reading the two-byte portion is sufficient to get a rough idea
of data contents. If particular data looks interesting, more and more data can
be read to improve the precision. This can easily be applied uniformly by each
process without coordinating. A more dynamic approach would be to apply a
wavelet compression~\cite{wavelet} approach that generates different compressed
data sizes at a fixed error rate depending on the data entropy. This is still a
uniform operator, but the resulting data can be of radically different sizes.

@@

The metadata system has a daunting challenge. Traditional metadata systems for
both distributed and parallel file systems only have to deal with a single name
space and storage hierarchy. While hierarichal storage management~\cite{hsm}
systems offer a way to show a single name space, the reality is that the tier
closest to the user is a cache for the other tiers with data migrating as
necessary to support user interaction. SIRIUS is working in a fundamentally
different way. Each tier is treated as a first class citizen with direct
client access possible. Where data is actually retrieved from is where it is
actually stored rather than forcing data migration and caching. Systems like
Sirocco~\cite{sirocco} offer a way to support this model directly and are a
fundamental part of our system.

@@

\section{Acknowledgements}

\includegraphics[scale=0.07]{logos/doe_logo}
\includegraphics[scale=0.30]{logos/snl_logo}
\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

\bibliographystyle{abbrv}
\bibliography{paper}

\vfill\eject

>>>>

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix,epsfig,endnotes,url}
\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}

%for single author (just remove % characters)
\author{
{\rm Your N.\ Here}\\
Your Institution
\and
{\rm Second Name}\\
Second Institution
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

%\author{
%\alignauthor {Jay Lofstead\footnotemark[1], Carlos Maltzahn\footnotemark[2], Scott Klasky\footnotemark[3], Hasan Abbasi\footnotemark[3], Qing Liu\footnotemark[3]\\
%Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
%\footnotemark[1]~~Sandia National Laboratories
%\footnotemark[2]~~University of California\\
%\footnotemark[3]~~Oak Ridge National Laboratory
%\footnotemark[4]~~Rutgers University
%\footnotemark[5]~~Brown University\\
%}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}

\subsection*{Abstract}
The longest existing parallel I/O APIs, HDF5, PnetCDF, and NetCDF, are focused
on treating the entire output from a paralell application as a single unit.
These libraries work to gather the data from all processes and arrange it such
that it is organized much like as if it were written from a single process.
This approach, in particular the two-phase collective I/O with data sieving
portion, has proven difficult to scale for 3-D domain decompositions at scale.
ADIOS shifted the focus from treating the entire application output as a single
entity to treating the output from each writer as a self-contained entity. This
eliminated the need for the problematic two-phase collective I/O and has
demonstrated good scalability well into the petascale era. Another shift is
required to address the needs of exascale systems. New system I/O bandwidth is
not keeping pace with the compute acceleration leading to accute bottlenecks.
While ``burst buffer'' technology will help address the performance gap, it
does not address the problem completely. This paper describes a new approach
to not just an I/O library, but also the underlying storage infrastructure to
address the needs of exascacle applications cognizant of projected exascale
platform characteristics and in the context of current industry trends.

@@

The first generation of standardized APIs for scientific computing IO focused
on serial output from a single process to a distributed or local file system.
These systems still exist in the form of serial HDF5~\cite{chilan:2006:hdf5-performance} and the
end-of-life NetCDF 3.x~\cite{netcdf3}. The key feature of these APIs was the
logical data model that eliminated any parallel data decomoposition artifacts
from the actual data storage. This offered consistent performance and regular
data access patterns.

The second generation is really an evolution of the first with HDF-5 adding
parallel capabilities and NetCDF bifurcating into NetCDF4~\cite{rew:2011:netcdf-guide} and
PnetCDF~\cite{torrellas:2003:pnetcdf} with PnetCDF almost completely maintaining the data
model and NetCDF4 shifting to use HDF5 underneath. The other improvement is
incorporating knowledge of an underlying parallel file system to support large
files written in unision from many processes.

The third generation is a reaction to the second based on recognizing that
maintaining the logical data model in physical form became inordinately
expensive and offered a performance penalty in most cases when running at
scale. Libraries like ADIOS~\cite{lofstead:2009:adaptable} shifted from writing complete
variables from distributed processes as a single entity to treating all data
from a single source to be treated as a unit. This eliminated the data
reorganization at the cost of more complicated reading patterns. In spite of
this extra complication, the performance is generally
better~\cite{lofstead:2011:six-degrees}.

@@

\noindent\textbf{Refactoring} may be a uniform operation across an
entire variable. For example, splitting a double into three parts--two byes,
four bytes, and two bytes, yields a two-byte value that is the majority of the
data precision. The next four bytes would be most of the rest of the data
precision. The last two bytes are the remainder of the precision. For quick
overviews, just reading the two-byte portion is sufficient to get a rough idea
of data contents. If particular data looks interesting, more and more data can
be read to improve the precision. This can easily be applied uniformly by each
process without coordinating. A more dynamic approach would be to apply a
wavelet compression~\cite{klappenecker:1995:wavelet} approach that generates different compressed
data sizes at a fixed error rate depending on the data entropy. This is still a
uniform operator, but the resulting data can be of radically different sizes.

@@

The metadata system has a daunting challenge. Traditional metadata systems for
both distributed and parallel file systems only have to deal with a single name
space and storage hierarchy. While hierarichal storage management~\cite{blaze:1992:hsm}
systems offer a way to show a single name space, the reality is that the tier
closest to the user is a cache for the other tiers with data migrating as
necessary to support user interaction. SIRIUS is working in a fundamentally
different way. Each tier is treated as a first class citizen with direct
client access possible. Where data is actually retrieved from is where it is
actually stored rather than forcing data migration and caching. Systems like
Sirocco~\cite{curry:2015:sirocco} offer a way to support this model directly and are a
fundamental part of our system.

@@

%\subsection{How to Build Your Paper}
%
%You have to run {\tt latex} once to prepare your references for
%munging.  Then run {\tt bibtex} to build your bibliography metadata.
%Then run {\tt latex} twice to ensure all references have been resolved.
%If your source file is called {\tt usenixTemplate.tex} and your {\tt
%  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
%{\tt \small
%\begin{verbatim}
%latex usenixTemplate
%bibtex usenixTemplate
%latex usenixTemplate
%latex usenixTemplate
%\end{verbatim}
%}
%

\section{Acknowledgments}

%\includegraphics[scale=0.07]{logos/doe_logo}
%\includegraphics[scale=0.30]{logos/snl_logo}
%\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

{\footnotesize \bibliographystyle{acm}
\bibliography{paper}}

%\theendnotes
